### 2019-08-21 paper/paper.tex
Improved parameter section opening.

@@

\section*{Problem Definition}

We search for 24 optimization variables that minimize the Handling Quality
Metric defined in ~\cite{Hess2012}.

Our formulation exposes 24 optimization variables
$\mathbf{p}\in\mathbb{R}^{24}$ that are both reformulations of and additions to
the parameters of the benchmark bicycle~\cite{Meijaard2007}. We separate the
rider's inertial parameters from the rear frame of the bicycle for the purposes
of rider/bicycle relative configuration but do not introduce any more than the
benchmark's four rigid bodies. The variables are as follows;
%
\begin{itemize}
  \itemsep-0.25em
  \item wheelbase, trail, steer axis tilt, wheel radii
  \item somersault angle of the person (same direction as bicycle pitch)
  \item masses of the front frame, rear frame, front wheel, and rear wheel
  \item mass center locations of the front frame, rear frame, and rider
  \item principal central radii of gyration of front and rear frames
  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
\end{itemize}

We formulate bounds on the variables as well as equality and inequality
constraints among the variables to ensure a physically realizable bicycle.
Below the constraints are defined and grouped by the associated rigid body or
collection thereof:
%
\begin{description}

  \item[Vehicle] The resulting combination of the five rigid bodies.
    \begin{itemize}
      \itemsep0em
      \item The physical extents of the bicycle and rider must exist above the
        ground plane.
      \item Both bicycle and rider are symmetric about the sagittal
        plane.
      \item The mass of each bicycle rigid body is positive, greater than a
        minimum value, and the total mass is below a reasonably lift-able
        amount.
      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
        radius.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration, i.e. -1 g < acceleration < 1 g.
    \end{itemize}

@@

The constraint statements above translate into upper and lower bounds on 12 of
the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
expressed in the standard non-linear programming form.
%
\begin{equation}
  \begin{aligned}
    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
    & \text{subject to} & & \\
    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
  \end{aligned}
\end{equation}

\subsection*{Parameters}
%
We introduce a new parameterization of the linear Whipple-Carvallo bicycle
model. This parameterization differs from the benchmark
parameterization~\cite{Meijaard2007} in that the person and rear frame are
separate rigid bodies each with their on inertial parameters and that the
inertial parameters of all bodies are expressed in terms of principal central
radii of gyration so that the mass is decoupled from the inertia. We also
introduce two simple dimensional parameters that define the geometric extents
of the person. Table~\ref{tab:parameters} provides the parameters and the
initial values which are derived from.

% TODO : Create a table of the parameters and values.
\begin{table}
  \caption{Parameters and their default values.}
  \begin{tabular}{llll}
    \toprule
    Variable & Value & Units & Description \\
    \midrule
    $c$ &  0.068581 & \si{\meter} & Trail \\
    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
    Rear Wheel [R] & & \\
    \midrule
    $m_R$ &  3.11 & \si{\kilogram} & mass \\
    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
    Front Wheel [F] & & \\
    \midrule
    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
    $r_F$ &  0.34353 & \si{\meter} & Radius \\
    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
    $y_F$ & 0.0 & \si{\meter} & Mass center \\
    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
    \midrule
    $l_P$ &  1.7280 & \si{\meter} & Body length \\
    $w_P$ &  0.48300 & \si{\meter} & Body width \\
    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
    $y_P$ & 0.0 & \si{\meter} & Mass center \\
    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
    \midrule
    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
    $y_H$ & 0.0 & \si{\meter} & Mass center \\
    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
    \midrule
    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
    $y_D$ & 0 & \si{\meter} & Mass center \\
    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
    \bottomrule
  \end{tabular}
\end{table}

These parameters can be transformed into the benchmark parameterization
r adily, but not vice versa.

We search for the optimal values of a subset of 23 of the principal parameters.

\begin{align}
  \mathbf{p} =
    [w \quad
     c \quad
     \lambda \quad
     m_D \quad
     x_D \quad
     z_D \quad
     k_{Daa} \quad
     k_{Dbb} \quad
     \alpha_D \quad
     x_P \quad
     z_P \quad
     \alpha_P \quad
     m_H \quad
     x_H \quad
     z_H \quad
     k_{Haa} \quad
     k_{Hbb} \quad
     k_{Hyy} \quad
     \alpha_H \quad
     r_R \quad
     m_R \quad
     r_F \quad
     m_F]
\end{align}

\subsection*{Constraints}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure a physically realizable bicycle is obtained.

@@

\subsection*{Rear Frame [D]}

There are five free parameters associated with the rear frame's inertial
characteristics:

@@

with these additional parameters that have fixed values:

\begin{align}
  I_{Dxy} = I_{Dyz} = 0 \\
  y_D = 0 \\
  % 
  l_P = ? \si{m}\\
  w_P = ? \si{m} \\
  m_P = 83.5 \si{kg} \\
  y_P = 0 \si{\meter} \\
  I_{Pxx} = 10.98524 \si{\kg\meter^2} \\
  I_{Pyy} = 11.2571 \si{\kg\meter^2}\\
  I_{Pzz} = 2.26922 \si{\kg\meter^2}\\
  I_{Pxy} = I_{Pyz} = 0 \si{\kg\meter^2}\\
  I_{Pxz} = -1.70211 \si{\kg\meter^2}\\
  % 
  I_{Hxy} = I_{Hyz} = 0 \\
  y_H = 0 \\
  %
  I_{Rxy} = I_{Ryz} = I_{Rxz} = 0 \\
  %
  I_{Fxy} = I_{Fyz} = I_{Fxz} = 0 \\
\end{align}

with these conversions from our parameters to the benchmark parameters:

\begin{align}
  I_{Dxx} = m_D k_{Dxx}^2 \\
  I_{Dyy} = \sqrt{I_{Dxx}^2 + I_{Dzz}^2} \\
  I_{Dzz} = m_D k_{Dzz}^2 \\
  I_{Dxz} = f(I_{Dxx}, I_{Dzz}, \alpha_D)  \\
  %
  m_B = m_D + m_P \\
  x_B = \frac{m_D x_D + m_P x_P}{m_B} \\
  z_B = \frac{m_D z_D + m_P z_P}{m_B} \\
  I_{Bxx} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Byy} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Bzz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  I_{Bxy} = 0 \\
  I_{Byz} = 0 \\
  I_{Bxz} = f(m_B, I_P, I_D, x_D, z_D, x_P, z_P, x_B, z_B) \\
  % 
  I_{Hxx} = m_H k_{Hxx}^2 \\
  I_{Hyy} = m_H k_{Hyy}^2\\
  I_{Hzz} = m_H k_{Hzz}^2 \\
  I_{Hxz} = f(I_{Hxx}, I_{Hzz}, \alpha_H)  \\
  %
  I_{Ryy} = m_R r_R^2 \\
  I_{Rxx} = I_{Ryy} / 2 \\
  I_{Rzz} = I_{Ryy} / 2 \\
  %
  I_{Fyy} = m_F r_F^2 \\
  I_{Fxx} = I_{Fyy} / 2 \\
  I_{Fzz} = I_{Fyy} / 2 \\
  %
  m_T = m_H + m_B + m_F + m_R \\
  x_T = \frac{m_H x_H + m_B x_B + m_F w}{m_T} \\
  z_T = \frac{m_H z_H + m_B z_B + m_F r_F + m_R r_R}{m_T} 
\end{align}

with these parameter bounds:

\begin{align}
  m_R > 0 \\
  m_F > 0 \\
  m_H > 0 \\
  m_D > 0 \\
  r_R > r_{min} \\
  r_F > r_{min} \\
  z_H < 0 \\
  z_D < 0 \\
  -\frac{\pi}{2} \leq \alpha_H \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \alpha_P \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \alpha_D \leq \frac{\pi}{2} \\
  -\frac{\pi}{2} \leq \lambda \leq \frac{\pi}{2}
\end{align}

and these constraints:

\begin{align}
  I_{Hyy} \leq \sqrt{I_{Hxx}^2 + I_{Hzz}^2} \\
  z_P + l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P + w_P / 2 \sin{\alpha_P} & \leq 0 \\
  z_P - l_P / 2 \cos{\alpha_P} & \leq 0 \\
  z_P - w_P / 2 \sin{\alpha_P} & \leq 0 \\
  w > r_F + r_R \\
  m_{Tmin} + m_P \leq m_T \leq 25 \si{kg} + m_P \\
  \frac{x_T}{|z_T|} < \frac{1}{4} \\
  \frac{x_T - w}{|z_T|} < -1 \\
  real(\mathbf{s}) < 0
\end{align}

>>>>

% - Abstract
% - Introduction
% - Problem Definition
%   - Optimization
%   - Parameterization
%   - Bounds
%   - Constraints
% - Solutions
%   - Parameters
%   - Eigenvalues
%   - HQM
% - Discussion and Conclusion

@@

\section*{Problem Definition}
%
We search for 24 optimization variables that minimize the Handling Quality
Metric defined in~\cite{Hess2012}.

\subsection*{Optimization}
%
The constraint statements above translate into upper and lower bounds on 12 of
the 24 variables $\mathbf{p}^U,\mathbf{p}^L$ and 14 linear and 21 nonlinear
constraint $\mathbf{g}(\mathbf{p})$ among the variables and thus can be
expressed in the standard non-linear programming form.
%
\begin{equation}
  \begin{aligned}
    & \underset{\mathbf{p}}{\text{minimize}} & & max(\textrm{HQM}(\mathbf{p})) \\
    & \text{subject to} & & \\
    & & & \mathbf{p}^L \leq \mathbf{p} \leq \mathbf{p}^U \\
    & & & \mathbf{g}^L \leq \mathbf{g}(\mathbf{p}) \leq \mathbf{g}^U \\
  \end{aligned}
\end{equation}

\subsection*{Bicycle Model Parameters}
%
Our problem formulation relies on a new parameterization that reflects both a
reformulation of and addition to the benchmark parameterization of the linear
Whipple-Carvallo bicycle model~\cite{Meijaard2007}. This parameterization
differs from the benchmark parameterization in that the person and rear frame
are treated as separate but fixed rigid bodies each with their on inertial
parameters and that the inertial parameters of all bodies are expressed in
terms of central principal radii of gyration so that the mass is decoupled from
the inertia. We also introduce two simple dimensional parameters that define
the geometric extents of the person. Table~\ref{tab:parameters} provides the
parameters and the initial values which are derived from the measurements of a
Batuvus Browser Bicycle and the rider, Jason, presented in~\cite{Moore2012}.
This allows the position and orientation of the rider to be specified
independtly of the rear frame but does not introduce any more than the
benchmark's four rigid bodies.

The variables are as follows:
%
\begin{itemize}
  \item wheelbase, trail, steer axis tilt, wheel radii
  \item somersault angle of the person (same direction as bicycle pitch)
  \item masses of the front frame, rear frame, front wheel, and rear wheel
  \item mass center locations of the front frame, rear frame, and rider
  \item principal central radii of gyration of front and rear frames
  \item principal inertia axis pitch angles of the front frame, rear frame, and rider
\end{itemize}

\begin{table}
  \caption{Parameters and their default values.}
  \begin{tabular}{llll}
    \toprule
    Variable & Value & Units & Description \\
    \midrule
    $c$ &  0.068581 & \si{\meter} & Trail \\
    $w$ &  1.1210 & \si{\meter} & Wheelbase \\
    $\lambda$ &  0.39968 & \si{\radian} & Steer axis tilt \\
    $g$ &  9.81 & \si{\meter\per\second\squared} & Acceleration due to gravity \\
    $v$ &  3.0 & \si{\meter\per\second} & Forward speed \\
    Rear Wheel [R] & & \\
    \midrule
    $m_R$ &  3.11 & \si{\kilogram} & mass \\
    $r_R$ &  0.34096 & \si{\meter} & Rear wheel radius \\
    $x_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $y_R$ & 0 & \si{\meter} & Rear wheel mass center \\
    $z_R$ & -0.34096 & \si{\meter} & Rear wheel mass center \\
    $k_{Raa}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration \\
    $k_{Rbb}$ &  0.17050 & \si{\meter} & Rear wheel central principal radii of gyration  \\
    $k_{Ryy}$ &  0.22136 & \si{\meter} & Rear wheel central principal radii of gyration \\
    Front Wheel [F] & & \\
    \midrule
    $m_F$ &  2.02 & \si{\kilogram} & Mass \\
    $r_F$ &  0.34353 & \si{\meter} & Radius \\
    $x_F$ &  1.1210 & \si{\meter} & Mass center \\
    $y_F$ & 0.0 & \si{\meter} & Mass center \\
    $z_F$ & -0.34353 & \si{\meter} & Mass center \\
    $k_{Faa}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fbb}$ &  0.20917 & \si{\meter} & Central principal radius of gyration \\
    $k_{Fyy}$ &  0.27179 & \si{\meter} & Central principal radius of gyration \\
    Person [P] & & \\
    \midrule
    $l_P$ &  1.7280 & \si{\meter} & Body length \\
    $w_P$ &  0.48300 & \si{\meter} & Body width \\
    $m_P$ &  83.500 & \si{\kilogram} & Mass \\
    $x_P$ &  0.31577 & \si{\meter} & Mass center \\
    $y_P$ & 0.0 & \si{\meter} & Mass center \\
    $z_P$ & -1.0990 & \si{\meter} & Mass center \\
    $k_{Paa}$ &  0.36797 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pbb}$ &  0.15276 & \si{\meter} & Central principal radius of gyration \\
    $k_{Pyy}$ &  0.36717 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_P$ & 0.18618 & \si{\radian} & Principal axis angle \\
    Front Frame [H] & & \\
    \midrule
    $m_H$ & 3.2200 & \si{\kilogram} & Mass \\
    $x_H$ & 0.86695 & \si{\meter} & Mass center \\
    $y_H$ & 0.0 & \si{\meter} & Mass center \\
    $z_H$ & -0.74824 & \si{\meter} & Mass center \\
    $k_{Haa}$ & 0.29556 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hbb}$ & 0.14493 & \si{\meter} & Central principal radius of gyration \\
    $k_{Hyy}$ & 0.27630 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_H$ & 0.36995 & \si{\radian} & Principal axis angle \\
    Rear Frame [D] & & \\
    \midrule
    $m_D$ &  9.8600 & \si{\kilogram} & Mass \\
    $x_D$ &  0.27595 & \si{\meter} & Mass center \\
    $y_D$ & 0 & \si{\meter} & Mass center \\
    $z_D$ & -0.53784 & \si{\meter} & Mass center \\
    $k_{Daa}$ &  0.28587 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dbb}$ &  0.22079 & \si{\meter} & Central principal radius of gyration \\
    $k_{Dyy}$ &  0.36539 & \si{\meter} & Central principal radius of gyration \\
    $\alpha_D$ &  1.1722 & \si{\radian} & Principal axis angle \\
    \bottomrule
  \end{tabular}
\end{table}

$\mathbf{p}\in\mathbb{R}^{24}$These parameters can be transformed into the benchmark parameterization
r adily, but not vice versa.
We search for the optimal values of a subset of 23 of the principal parameters.

\begin{align}
  \mathbf{p} = &
    [w \quad
     c \quad
     \lambda \quad
     m_D \quad
     x_D \quad
     z_D \quad
     k_{Daa} \quad
     k_{Dbb} \quad
     \alpha_D \quad
     x_P \quad
     z_P \quad
     \alpha_P \quad \\
 &    m_H \quad
     x_H \quad
     z_H \quad
     k_{Haa} \quad
     k_{Hbb} \quad
     k_{Hyy} \quad
     \alpha_H \quad
     r_R \quad
     m_R \quad
     r_F \quad
     m_F]
\end{align}

\subsection*{Bounds and Constraints}
%
The optimal principal parameters are subject to a set of constraints designed
to ensure a physically realizable bicycle is obtained.
We formulate bounds on the variables as well as equality and inequality
constraints among the variables to ensure a physically realizable bicycle.
Below the constraints are defined and grouped by the associated rigid body or
collection thereof:
%
\begin{description}
  \item[Vehicle] The resulting combination of the five rigid bodies.
    \begin{itemize}
      \itemsep0em
      \item The physical extents of the bicycle and rider must exist above the
        ground plane.
      \item Both bicycle and rider are symmetric about the sagittal
        plane.
      \item The mass of each bicycle rigid body is positive, greater than a
        minimum value, and the total mass is below a reasonably lift-able
        amount.
      \item The wheels cannot overlap, i.e. wheelbase > front radius + rear
        radius.
      \item The bicycle cannot topple forward during hard braking or backward
        during hard acceleration, i.e. -1 g < acceleration < 1 g.
    \end{itemize}

@@

@@

\section*{Rigid Body Definitions}

\subsection*{Rear Frame [D]}
%
There are five free parameters associated with the rear frame's inertial
characteristics:

@@
